Finite Complete Suites for CSP Refinement Testing

### 2018-11-19 scp-journal/article/finitecomplete-failures.tex
Section 4 revised

\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
after having run through trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$ by
assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$. Since we assume that
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that $[G(P)/s]^0$ is
{\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
$e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that
$\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be
entered, and the test stops after $\epass$. Otherwise, if $\minhits(G(P)/s)
\neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$ contains only
elements that are contained in some minimal acceptance of $P/s$, and all
these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is a subset of
$[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a hitting set of
$\minaccs(Q/s)$, also $H$ is not a hitting set of $\minaccs(Q/s)$. Now this
is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by
assumption, so $H$ should also be a (not necessarily minimal) hitting set in
$\minaccs(Q/s)$. This proves that the test execution cannot block  at branch
(\ref{eq:ufc}) without being able to pass the test by entering branch
(\ref{eq:ufb}).

\paragraph{Step~2.2} Now suppose that the test execution blocks at
branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq \trc(P)$
with $\#s = j$.
From Lemma~\ref{lemma:ufproperties} we know
that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
$U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
\varnothing$.
If $\minhits(n) \neq \varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$
blocks at branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$
which is not a hitting set of $\minaccs(Q/s)$. Again, we can apply Lemma~\ref{lemma:hseta}
to conclude that this contradicts the assumption that $Q\ conf\ P$. As a consequence,
the test execution can never deadlock at branch (\ref{eq:ufd}) without entering
branch (\ref{eq:ufb}) and passing the test.

Note that the line of reasoning in this proof requires that $Q$ is free of
livelocks, because otherwise a $\epass$ event might not become visible,
due to unbounded sequences of hidden events performed by $Q$.
\xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the
reference process $P$ and the SUT process $Q$: any trace
$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
through this product graph. Any error can be detected after running through
such a trace and then either observing an event outside $[P/s]^0$ (this is
the violation described by Case~1) or identifying an illegal acceptance $A_Q$
(as specified in Case~2). It is not guaranteed, however, that trace $s$ is
short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
pq$. So, it has to be shown that for any $s$ leading to an error situation
specified by Case~1 or Case~2, there exists a trace $u$ of maximal length
$pq-1$ leading to the same error.

@@

\medskip
\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
product graph $G$, but we do not know its length. Again, by applying
Lemma~\ref{lemma:reachproduc}, we know that the state $G/s$ can be reached by
a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider
the test $U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because
of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
(\ref{eq:ufc}) until the trace $u$ has been completely processed.
 $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
only:~(\ref{eq:accsnotcontained}) implies that $P/s$ has at least one
non-empty minimal acceptance. By (\ref{eq:minhitminaccempty}) this is
equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and we
observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and
therefore fails for the execution where $Q/u$ selects the minimal acceptance
$A_Q$ as specified in (\ref{eq:accsnotcontained}) and $U_F(\#u,\#u,G(P)/u)$
selects a minimal hitting set $H\in\minhits(P/u)$ that has an empty
intersection with $A_Q$. The existence of such an $H$ is guaranteed because
of Lemma~\ref{lemma:hseta}. As a consequence, $(Q\parallel[\Sigma]U_F(\#
u))/u$ cannot produce the $\epass$ event in this execution; this means that
the test fails. The complete testing assumption guarantees that this
execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed
sufficiently often. This concludes the proof. \xbox
\end{proof}
%
Our notion of tests can be specialised to deal with traces refinement, as we
explain in Section~\ref{sec:finitecomplete}.
The next section presents and example.

>>>>

\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
after having run through a trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$
by assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$. Since we assume that
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that $[G(P)/s]^0$ is
{\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
$e\in [G(P)/s]^0$ would be accepted by $Q/s$ and the test execution would not
deadlock. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch
(\ref{eq:ufb}) can be entered, and the test stops after $\epass$. Otherwise,
if $\minhits(G(P)/s) \neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$
contains only elements that are contained in some minimal acceptance of
$P/s$, and all these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is
a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a
hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of
$\minaccs(Q/s)$. Now this is a contradiction to Lemma~\ref{lemma:hseta},
since $Q\ conf\ P$ by assumption, so $H$ should also be a (not necessarily
minimal) hitting set in $\minaccs(Q/s)$. This proves that the test execution
cannot block  at branch (\ref{eq:ufc}) without being able to pass the test by
entering branch (\ref{eq:ufb}).

\paragraph{Step~2.2} Suppose that the execution blocks at
branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq
\trc(P)$ with $\#s = j$. From Lemma~\ref{lemma:ufproperties} we know that
$U_F(j)/s = U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so $\minhits(n) =
\minhits(P/s)$. Branch (\ref{eq:ufb}) of $U_F(j,k,n)$ leads always to a PASS
verdict and is taken if $\minhits(n) = \varnothing$. If $\minhits(n) \neq
\varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$ blocks at
branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$ that
is not a hitting set of $\minaccs(Q/s)$. Again, by Lemma~\ref{lemma:hseta},
this contradicts the assumption that $Q\ conf\ P$. As a
consequence, the test execution can never deadlock at branch (\ref{eq:ufd})
without entering branch (\ref{eq:ufb}) and passing the test.

\bigskip \noindent%
Note that the line of reasoning in this proof requires that $Q$ is free of
livelocks, because otherwise a $\epass$ event might not become visible, due
to unbounded sequences of hidden events performed by $Q$. \xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a
    minimal acceptance $A_Q$ of $\minaccs(Q/s)$, such that (see
    Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}))
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the
reference process $P$ and the SUT process $Q$: any trace
$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
through this product graph. Any error can be detected after running through
such a trace and then either observing an event outside $[P/s]^0$ (the
violation described by Case~1) or identifying an illegal acceptance $A_Q$ (as
in Case~2). It is not guaranteed, however, that $s$ is short enough to be
executed by one of the test cases $U_F(j)$ with $0\le j < pq$. So, it has to
be shown that for any $s$ leading to an error situation, there exists a trace
$u$ of maximal length $pq-1$ leading to the same error.

@@

\medskip
\noindent {\bf Case~2.} We note again that $s$ is a trace of the product
graph $G$, but we do not know its length. Again, by
Lemma~\ref{lemma:reachproduc}, the state $G/s$ can be reached by a trace
$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider the test
$U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of
Lemma~\ref{lemma:ufproperties}. $U_F(\#u)$ can always perform branch
(\ref{eq:ufc}) until the trace $u$ has been completely processed.
 $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
only:~(\ref{eq:accsnotcontained}) implies that $P/s$ has at least one
non-empty minimal acceptance. By (\ref{eq:minhitminaccempty}) this is
equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and we
observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and
therefore fails for the execution where $Q/u$ selects the minimal acceptance
$A_Q$ as specified in (\ref{eq:accsnotcontained}) and $U_F(\#u,\#u,G(P)/u)$
selects a minimal hitting set $H\in\minhits(P/u)$ that has an empty
intersection with $A_Q$. The existence of such an $H$ is guaranteed because
of Lemma~\ref{lemma:hseta}. As a consequence, $(Q\parallel[\Sigma]U_F(\#
u))/u$ cannot produce the $\epass$ event in this execution; this means that
the test fails. The complete testing assumption guarantees that this
execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed
sufficiently often. This concludes the proof. \xbox
\end{proof}
%
Our notion of test can be specialised to deal with traces refinement~(see
Section~\ref{sec:finitecomplete}). We next present an example.
